From Questions to Queries – Enhancing the Curt System by a Theory and Implementation of Embedded WH-Questions

### 2008-08-22 thesis/paper.tex
more work.

%fixme: to be formalized
Where we use ``natural language'', usually the English language will be treated
in practice, though it would be nice to come up with treatments of other
languages. We abstract away from pragmatic affairs and concern ourselves with
mostly the purely semantic/logic domain, so this should be feasible to achieve.
\subsection*{Acknowledgements}

@@

% fixme: formalize
A brief overview of the premisses for the theory (which itself lays ground to
the implementation).

@@

\pn{Curt}, a system written in \prol that is able to interpret basic utterances
in natural language and store them in expressions of first order logic, is presented
in \cite{blackburnbos:cl1}. While relatively limited when compared to other
systems presented in \ref{sec:comparision} in
terms of lexical and grammatical coverage it is able to perform basic checks for
\emph{consistency} and \emph{informativity} of its input\footnotemark, within
the bounds of the tools it uses.

@@

>>>>

Growing efforts have thus been made to provide a technology that is (going to
be) able to extract logical formulae from sentences of ordinary language and
establish relations between objects of its logical representation. In addition,
providing a means to compute with said representations is will eventually enable
the machine-processing of meaning.  In doing so, the ultimate goal may be to
enable a man-made system to \emph{understand} what a humans communicate.

Exploration of this  interesting field of linguistics has shown that
the notion of \emph{understanding} is far more than a linguistic concern, but
involves at least findings in the domain of logics, (cognitive) psychology and
probably even philosophy.

Development of exhaustive systems in this area appears to be a daunting task of
quite some scope. Therefore most approaches have been centered on providing
coverage of at least a fragment of natural language. Many such systems exist and
they prove to be increasingly successful. Some of them will be discussed here
and presenting an extension to one system in particular is the main matter of
this Bachelor's thesis.

In doing so, the present work will aim to provide the \curt system of
\cite{blackburnbos:cl1} with the mechanics of dealing with more than one
partition of its logical domain, viz. the ability to handle more than one
\term{possible} or \term{accessible world}. As a ground for development of this
work, the treatment of \term{embedded questions} was chosen as one that
naturally demands a possibility to account for more than one division in a
system's \emph{understanding} or representation of a universe of objects and
relations.

As

The first section gives a conceptual overview, introducing the \curt
system and comparing it to other similar systems in existence. It then goes on
to explain the target domain of natural language the implementation has to
account for.
The next section concerns itself with a more detailed analysis of the theoretic
assumptions of this implementation and justifies some choices and restrictions
that had to be made because of the limited scope of this analysis. While the
third section will display and clarify the actual implementation on top of the
\curt system, the last section will discuss the results of this work and also
give a prospect on possible further work on the subject.

\subsection*{Acknowledgements}

Frank... who else?

@@

The basis of

%fixme: to be formalized
Where we use ``natural language'', usually the English language will be treated
in practice, though it would be nice to come up with treatments of other
languages. We abstract away from pragmatic affairs and concern ourselves with
mostly the purely semantic/logic domain, so this should be feasible to achieve.

A brief overview of the premisses for the theory (which itself lays ground to
the implementation).

@@

\pn{Curt}, a system written in \prol that is able to interpret basic utterances
in natural language and store them in expressions of first order logic, is
presented in \cite[chapter 6]{blackburnbos:cl1}. While relatively limited when
compared to other systems presented in \ref{sec:comparision} in terms of lexical
and grammatical coverage, it provides a sound basis for extended work because of
its coherent design and relative simplicity.

Besides parsing and translating utterances into logic, \curt is able to perform
basic checks for \emph{consistency} and \emph{informativity} on the resulting
formulae -- within the bounds of the tools it uses.

@@

To quote \cite{blackburnbos:cl1}:

\begin{quote}
  Questions, unlike assertions, don't have truth-values, so it would be
  misleading to represent them as ordinary formulas.
\end{quote}

It is important to note that this claim is not undisputed and
\cite{karttunen:1977}, argues against it. The present approach, however will
base itself on this assumption and extend it by more formal approaches than the
one given by Blackburn \& Bos. Specifically, the work by J. Groenendijk and M. Stokhof
will lay ground to the implementation. But in order to realise their formal
discussions computationally, modifications will have to be made, and even
compromises.
